---
title: "Initial Korean covid 19"
author: "ediaz16"
date: "Spring 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

# Introduction

$H_0:$ There is no difference in death rate based on sex

$H_A:$ There is a difference in the death rate based on Sex



# Method
Variables for this study are SEX and result.

We want to know if results depends on Sex  
result is a categorical variable, 
Sex is also a categorical variable

So we use Cat~ Cat helper file 

# Results
## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result+ Sex,data=kRdata, type="percent")

```

This chart shows that male death is higher on average being 60% and female average below 40%
  





  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

This table shows the average male deaths are higher than female deaths at 64 deaths vs female death 36.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test shows P value for table 1 is .001 show there is a significant difference this means there is not a difference in death rate based on Sex.

The p value on the Fisher Exact test is 0.0000000009.

The odds ratio is 2.3:1 this means 



## Conclusions

The original question was was "is there a difference in death rate in the Korean data set based on Sex?

The plot indicaed ...

The numericaL result was that ....

and we found a significant death rate differnce due to Sex was not significant 